---
title: "Using treevalues to perform inference on CART Trees"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{inference_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(treevalues)
```


# Introduction

All of the functions in this package assume that the user began by building a regression tree using \texttt{rpart}, an implementation of CART (Breiman et al., 1984). Throughout this tutorial, we work with an example tree fit to the Box Lunch Study dataset, which is the example dataset provided by CITATION in the \texttt{visTree} package. 

- important points = model=TRUE, fixed CP and relationship to alpha,

## Identifying nodes in an rpart tree

Unfortunately, rpart's method of identifying nodes in a tree is a little confusing. We use their method for consistencym, but we explain it here because its strange. In the last section of this paper, we show how their is some wiggle room :). 

# Inference on a single terminal node

Really easy- just get the node number and then run ``



# Inference on a pair of sibling terminal nodes



# Inference on a Full Tree

For convenience, we provide a method that returns a matrix with inference on a whole tree. And a method for plotting your enture tree with p-values and confidence intervals labeled. 

**challenge for anna: make the plotting method have more arguements! User can decide things like what is labeled and font size and stuff). 


# Mixing and matching: inference on non-terminal nodes or on permuted branches. 

As 



